---
title: "P8106 HW3"
author: "Shihui Zhu"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhead[R]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
--- 

\newpage

```{r setup, include=FALSE}
# This chunk loads all the packages used in this homework
library(caret)
library(MASS)
library(mlbench)
library(pROC)
library(klaR)
library(ggplot2)
library(tidyverse)

# General figure set up
knitr::opts_chunk$set(
  # hide warning messages
  warning = FALSE
)
```

## Car Gas Mileage

Data Description: 

Dependent: 
* mpg_cat: whether the miles per gallon of a car is high or low

Predictors:
* cylinders: Number of cylinders between 4 and 8
* displacement: Engine displacement (cu. inches)
* horsepower: Engine horsepower
* weight: Vehicle weight (lbs.)
* acceleration: Time to accelerate from 0 to 60 mph (sec.)
* year: Model year (modulo 100)
* origin: Origin of car (1. American, 2. European, 3. Japanese)

### Read in Data and Split training and testing set
```{r input}
cars <- read_csv("auto.csv") %>% janitor::clean_names() %>% na.omit()
# training data (70%) and test data (30%)
set.seed(1)
rowTrain <- createDataPartition(y = cars$mpg_cat,
                                p = 0.7,
                                list = FALSE)
```

### (a) Produce some graphical or numerical summaries of the data

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
